State and Parameter Estimation for Spatio-Temporal Neural Fields

### 2010-05-13 ltx/Journal of Neuroengineering/StateAndParameterEstimationNeuralFields_JoNE.tex
finished going through the body of the paper, making loads of little edits and correcting typos as I see 'em. Got up to the end of Section 7, haven't looked at the Appendices.

The unscented Rauch Tung Striebel Smoother (URTSS)~\cite{Sarkka2010} is used for the state estimation. The URTSS incorporates an unscented Kalman Filter (UKF)~\cite{Julier1997, Merwe2003} in a forward iteration to estimate posterior states, $\hat{\mathbf x}_t^{f}$, followed by a backward iteration to compute the smoothed state estimates, $\hat{\mathbf x}_t^{b}$. The first and the second order moments of the predicted state are captured by propagating the so-called sigma points through the state equation. The sigma points, $\mathcal X_i$, are calculated using the unscented transform as follows
\begin{equation}\label{eq:sigmapoints1}
	\mathcal X_{0}=\bar x 
	\mathcal X_{i}=\bar x+(\sqrt{( L + \lambda)\mathbf P_x})_i \quad i=1, \dots, L 
\begin{equation}\label{eq:sigmapoints2}
	\mathcal X_{i}=\bar x-(\sqrt{( L + \lambda)\mathbf P_x})_{i- L} \quad i= L+1, \dots, 2 L 
where $\bar x$ represents either $\hat{\mathbf x}_t^{f}$ or $\hat{\mathbf x}_t^{b}$, $\mathbf{P}_x$ is the corresponding covariance matrix from the filtering or smoothing, $(\sqrt{( L + \lambda)\mathbf P_x})_i$ is the $i$th column of the weighted matrix square root of $\mathbf P_x$ and $L$ is the dimension of the state space. The total number of sigma points is $2L+1$. The scaling parameter $\lambda$ is defined as 
\begin{equation}\label{eq:sigmapoints3}
	\lambda=\alpha^2( L+\kappa) - L 
where $\alpha$ (determines spread of sigma points), $\beta$ (incorporate prior knowledge of the distribution of $\mathbf{x}$) and $\kappa$ (scaling parameter) are constants (see ~\cite{Haykin2001} for more details). The sigma vectors are propagated through the system equations and weighted to form the predicted mean and covariance. The weights are calculated by 
	\mathbf W_0^{(m)}=\frac{\lambda}{ L+\lambda} 
	\mathbf W_0^{(c)}=\frac{\lambda}{ L+\lambda}+(1-\alpha^2+\beta) 
	\mathbf W_i^{(m)}=\mathbf W_i^{(c)}=\frac{1}{2( L+\lambda)} \quad i=1, \dots, 2L. 
Since the observation equation is linear (equation~\ref{ObservationEquation}) the standard Kalman Filter update equations are  used to correct the predicted states. The state estimates from the forward filtering are used to form a new set of sigma points for the smoother, as described above. To compute the smoother gain, the predicted cross-covariance matrix of the states is also calculated. A summary of the URTSS procedure is given in Algorithm~\ref{UKFAlgorithm}. It should be noted that as the disturbance, $e_t\left(\mathbf{r}\right)$ and the measurement noise $ \boldsymbol{\varepsilon}_t$ are additive, therefore the additive form of the URTS is used, otherwise the state vector should be augmented with the noise terms. The additive form reduces the dimension and the number of sigma points used in the algorithm.
\begin{algorithm}
	\caption{The Unscented RTS Smoother}\label{UKFAlgorithm} 
	\begin{algorithmic}[1] 
		\State Forward initialisation 
		 \hat{\mathbf x}_0, \mathbf P_0 
		\State Forward iteration: for $t \in \left\lbrace 0,\cdots, T\right\rbrace $,
		calculate the sigma points $\mathcal X_{i,t}^f$ using \ref{eq:sigmapoints1}-\ref{eq:sigmapoints3} and propogate them through equation \ref{eq:QmatrixForSigmapoints}
% 		\begin{small}
			\mathcal X_{i,t+1}^{f-}=Q(\mathcal X_{i,t}^f,u_t) 
% 		\end{small}
		calculate the predicted state and the predicted covariance matrix
			\hat{\mathbf x}_{t+1}^{f-}=\sum_{i=0}^{2L} W_i^{(m)}\mathcal X_{i,t+1}^{f-} 
			\mathbf P_{t +1}^{f-}=\sum_{i=0}^{2L} W_i^{(c)}(\mathcal X_{i,t+1}^{f-}-\hat{\mathbf x}_{t +1}^{f-})(\mathcal X_{i,t+1}^{f-}-\hat{\mathbf x}_{t +1}^{f-})^\top+\boldsymbol \Sigma_e 
		compute the filter gain, the filtered state and the filtered covariance matrix using the standard Kalman Filter update equations
			\mathcal K_{t+1}=\mathbf P_{t +1}^{f-}\mathbf C ^\top(\mathbf C \mathbf P_{t +1}^{f-}\mathbf C ^\top+\boldsymbol \Sigma_v)^{-1} 
			\hat{\mathbf x}_{t+1}^{f}=\hat{\mathbf x}_{t+1}^{f-}+\mathcal K_{t+1}(\mathbf y_{t+1}-\mathbf C\hat{\mathbf x}_{t +1}^{f-}) 
			\mathbf P_{t+1}^f=(\mathbf I - \mathcal K_{t+1}\mathbf C)\mathbf P_{t +1}^{f-} 
		\State Backward initialisation 
			\mathbf P_T^b= \mathbf P_T^f, \quad \hat{\mathbf x}^b_T= \hat{\mathbf x}^f_T 
		\State Backward iteration: for $t \in \left\lbrace T-1, \cdots, 0 \right\rbrace $ calculate the sigma points $\mathcal X_{i,t}^b$ and propogate them through equation \ref{eq:QmatrixForSigmapoints}
			\mathcal X_{i,t+1}^{b-}=Q(\mathcal X_{i,t}^b,u_t) 
		 compute the predicted state, the predicted covariance matrix and the cross-covariance matrix
			\hat{\mathbf x}_{t+1}^{b-}=\sum_{i=0}^{2L} W_i^{(m)}\mathcal X_{i,t+1}^{b-} 
			\mathbf P_{t +1}^{b-}=\sum_{i=0}^{2L} W_i^{(c)}(\mathcal X_{i,t+1}^{b-}-\hat{\mathbf x}_{t +1}^{f})(\mathcal X_{i,t+1}^{b-}-\hat{\mathbf x}_{t +1}^{f})^\top+\boldsymbol \Sigma_e 
			\mathbf M_{t +1}=\sum_{i=0}^{2L} W_i^{(c)}(\mathcal X_{i,t}^{b-}-\hat{\mathbf x}_{t}^{f})(\mathcal X_{i,t+1}^{b-}-\hat{\mathbf x}_{t+1}^{f})^\top 
		 Compute the smoother gain, the smoothed state and the smoothed covariance matrix
			\mathbf S_t=\mathbf M_{t +1}\left[ \mathbf P_{t +1}^{b-}\right] ^{-1} 
			\hat{\mathbf x}_t^b=\hat{\mathbf x}_t^f+\mathbf S_t\left[\hat{\mathbf x}_{t+1}^{b}-\hat{\mathbf x}_{t+1}^{b-}\right] 
			\mathbf P_{t}^{b}=\mathbf P_{t}^{f}+\mathbf S_t\left[\mathbf P_{t+1}^{b}-\mathbf P_{t+1}^{b-} \right]\mathbf S_t^\top 
	\end{algorithmic}
\end{algorithm}

Although the system is nonlinear, the parameters of the system are linear with respect to the state. This is exploited by our procedure where the parameter estimation uses a Least Squares (LS) method that minimises the sum of the squared errors (of a predicted state update) with each new state estimate (see Appendix~\ref{LeastSquaresAppendix}).

@@

\section{Results}\label{ResultsSection} The neural field model derived in Section~\ref{NeuralModelSection} was used to generate the data for state and parameter estimation. All parameters for the model are given in Table~\ref{tab:Model Parameters}. An example of the simulated field can be seen in figure~\ref{fig:experimental design}~A. The Figure demonstrates the experimental design where the sensors were placed in the centre of the field (full width at half maximum). This design allowed for uncertainty around the edges of the sensor array, which would be typical from physical data, and alleviated problems associated with boundary conditions. Figure~\ref{fig:experimental design}~B. shows example data from the sensors of the model, and figure~\ref{fig:experimental design}~C. shows the power spectral density of the simulated time-series. The PSD shows the typical $1/f$ characteristics of intracranial EEG.

@@

\subsection{Spatial Frequency Analysis} In accordance with Section~\ref{SpectralAnalysisSection}, the spatial frequency of the field was used to confirm the spacing and width of the sensors was adequate to capture the significant dynamics of the field. Figure~\ref{fig:FieldFFT} (part a.) shows the spatial frequency of the field. The cut-off frequency, $\nu_c$, was taken to be \omg{??} ($-3$~dB point). From equation~\ref{eq:MinimumSensorDistance}, this yielded a minimum sensor separation of \omg{??}. This confirms our separation of \omg{??} was enough to prevent problems associated with aliasing.

The spatial frequency of the observations is shown in Figure~\ref{fig:FreqAnalysis} (part c.). The peak power of the spatial frequency of the observations was approximately \omg{$??$}~dB. The cut-off frequency (-3~dB point) was taken to correspond to  approximately \omg{$??$}~dB. This provided a conservative estimate of the spatial bandwidth. The oversampling parameter $\rho_b$ was set to $??$ giving a minimum distance between basis functions of $??$ to satisfy Shannon's sampling criterion (from equation~\ref{eq:BasisFunctionSeparation}). \todo{maybe from Parham's derivation... From equation~\ref{} the width was calculated to be... or...}. The width of the basis functions was chosen such that they had sufficient overlap to represent the dynamic field.

\begin{figure*}
	\begin{center}
\subfigure[][]{\includegraphics[scale=0.8]{./Graph/BasisChar.pdf}}
\subfigure[][]{\includegraphics[scale=0.8]{./Graph/SensorChar.pdf}}
\subfigure[][]{\includegraphics[width=1.8in]{./Graph/ObservationFrequencyResponse.pdf}}\\
	\end{center}
	\caption{a:Field basis function frequency response (projected to a 1D plot because it is symmetric); b:Sensor frequency response (projected to a 1D plot because it is symmetric); c: The average FFT of the observed field.} 
\label{fig:FreqAnalysis} 
\end{figure*}
Following this, the distance between basis function, $\Delta_b$, was set to $2.5$ and the basis function width, $\sigma_{\phi}^2$, was set to 3.6, 3.16 mm full width at half maximum. Analysis said 3.9 from Parham's derivation, but 3.6 still overlaps and allows more high frequency details to be represented.

\subsection{State and Parameter Estimates} To demonstrate the performance of the parameter and state estimation, $250$ realisations of the data were generated. Each realisation consisted of 200 ms of data and the estimation was applied to the final 200 ms, allowing the model to stabilise. The initial state and parameters were unknown to the estimator. The distribution of the parameters and the resulting kernel are shown in figure~\ref{fig:Parameters} and figure~\ref{fig:KernelEstimates} respectively. Although a wide distribution can be seen over the parameter estimates the ratio between them remans in a narrow range as shown in figure~\ref{fig:ParametersRatio}. The state estimates for one realization were evaluated by comparing the field reconstruction to the true field using the mean over time of the Root Mean Square Error (RMSE) and standard deviation of the RMSE giving $0.19\pm 0.01$. The RMSE of the field estimate over time is also shown in figure~\ref{fig:RMSE}. One snapshot for the original and the field estimate and their corresponding error are illustrated in figure~\ref{fig:FieldEstimate}.
\begin{figure*}
\subfigure[][]{\includegraphics{./Graph/KernelEstimate_skinny.pdf}}
\subfigure[][]{\includegraphics{./Graph/KernelEstimateFFT_skinny.pdf}}\\
\subfigure[][]{\includegraphics{./Graph/KernelEstimate_wide.pdf}}
\subfigure[][]{\includegraphics{./Graph/KernelEstimateFFT_wide.pdf}}\\
\subfigure[][]{\includegraphics{./Graph/KernelEstimateFFT_skinny_db.pdf}}

@@

\omg{FIRST DESCRIBE AN OVER ALL SUMMARY OF WHAT WE HAVE DONE AND WHAT IS NOVEL}
This paper presents a novel framework for model reduction to facilitate practical state and parameter estimation of neural fields. A basis function decomposition reduces the infinite dimensional neural field (continuous space), to a finite dimensional state-space model. The level of detail of the estimated field is related to the basis functions decomposition, thereby flexible framework is provided where a trade-off can be made between physiological detail and computational complexity. By forming a state-space model, the unscented Rauch Tung Striebel Smoother was implemented for state estimation. Using the sequence of state estimates, a least squares algorithm was implemented for parameter estimation.

\omg{DISCUSS RESULTS: FREQ ANALYSIS. USE THIS TO JUSTIFY KERNEL ERRORS.}
The frequency analysis and model selection tools presented in this paper specify the sensor and field basis function arrangement required to form a state-space model that is capable of capturing the dominant cortical dynamics. These tools can be also used to better design electrophysiological recording techniques to advance model-based neural parameter estimation in order to avoid spatial aliasing. The effect of the basis function decomposition will restrict the frequency content of the estimated field. This is evident in when comparing Figures~\ref{fig:FieldFFT} and~\ref{fig:EstimatedFieldFFT} where the estimated field is spatially band-limited. The spatial frequency content of the field is governed by disturbance covariance and the shape of the connectivity kernel, therefore some information will be lost due to the field decomposition. The implication of this loss of information is the high spatial frequency content of the connectivity kernel. This is demonstrated in Figure~\ref{fig:KernelEstimates} of the results where the frequency content of the estimated connectivity kernel is less than the true kernel.

\omg{DISCUSS THAT EVEN THOUGH THE KERNEL ESTIMATE IS NOT EXACT, IT IS USEFUL TO KNOW THE RATIO OF EXCITATION TO INHIBITION AND THAT RELATIVE MEASURES CAN BE ACCURATELY TRACKED}
Even though the high frequency content of the connectivity kernel is lost though the basis function decomposition, the dominant dynamics are still captured with the estimated reduced model. This is illustrated in~\ref{fig:FieldEstimate}. In addition, the ratio of the kernel parameters follows a tight distribution as seen in Figure~\ref{fig:ParametersRatio}. This indicates that the ratio of excitation to inhibition can be tracked with this approach.

@@

\omg{DISCUSS THE IMPLICATION OF ESTIMATING INTRACORTICAL CONNECTIVITY AND WHERE WE FIT WITHIN OTHER FUNCTIONAL CONNECTIVITY RESEARCH AREAS}
To the authors best knowledge, this is the first paper to propose an estimation framework for intracortical connectivity from electrophysiological data. The importance of this can be illustrated by considering the connectivity structure at this scale can not be estimated using diffusion tensor imaging due to the near isotropic diffusion or water in grey matter~\cite{Assaf2008}. In addition, other non-model based techniques for estimating functional connectivity such as Granger causality \cite{Hesse2003}, the direct transfer function \cite{Kaminski1991}, and partial directed coherence \cite{Sameshima1999} are restricted to the spatial scales of the recording electrodes. The model-based technique proposed in this paper estimates the connectivity structure independently of the recording electrode geometry. An advantage of the auto-regressive (AR) based methods is that they allow for anisotropic connectivity, which is present in cortico-cortical fibres. It should be noted, however, that the AR and the proposed model-based methods are complementary, where connectivity is estimated at different spatial scales. Therefore, we envisage by combining these methods a more realistic model can be acquired. A recent theoretical study~\cite{Jirsa2009} has demonstrated the importance of this multi-scale structure in generating the characteristic rhythms of the brain.

@@

\omg{DISCUSS SHORT-COMINGS AND FUTURE WORK}
In this paper the widths of the connectivity kernel basis functions are assumed to be known. In a future work we plan to extend this work to unknown widths. We hope to achieve this by using a sufficiently large set of connectivity basis functions within the physiological plausible limits, and letting estimation procedure take care of the weights. In addition, we plan to show in future work how the assumption homogeneity can be relaxed, where heterogeneity can be modelled by careful placement of field basis functions. For example, if cortical regions have a greater density of neural populations, then more basis functions can be used in these areas. In this scenario, model selection and basis function positioning can be achieved using instantaneous frequency analysis of the the neural field or alternative model selection tools based on information criteria. This is a topic for future research. Future work should also be targeted to include a more realistic synaptic response kernel, with finite rise and decay times and finite propagation times into the neural field equations.

>>>>

The unscented Rauch Tung Striebel Smoother (URTSS)~\cite{Sarkka2010} is used for the state estimation. The URTSS incorporates an unscented Kalman Filter (UKF)~\cite{Julier1997, Merwe2003} in a forward iteration to estimate posterior states, $\hat{\mathbf x}_t^{f}$, followed by a backward iteration to compute the smoothed state estimates, $\hat{\mathbf x}_t^{b}$. The first and the second order moments of the predicted state are captured by propagating the so-called sigma points through the state equation. The sigma points, $\mathcal X_i$, are calculated using the unscented transform as follows
\begin{equation}\label{eq:sigmapoints1}
	\mathcal X_{0}=\bar x 
\end{equation}
\begin{equation}
	\mathcal X_{i}=\bar x+(\sqrt{( L + \lambda)\mathbf P_x})_i \quad i=1, \dots, L 
\end{equation}
\begin{equation}\label{eq:sigmapoints2}
	\mathcal X_{i}=\bar x-(\sqrt{( L + \lambda)\mathbf P_x})_{i- L} \quad i= L+1, \dots, 2 L 
\end{equation}
where $\bar x$ represents either $\hat{\mathbf x}_t^{f}$ or $\hat{\mathbf x}_t^{b}$, $\mathbf{P}_x$ is the corresponding covariance matrix from the filtering or smoothing, $(\sqrt{( L + \lambda)\mathbf P_x})_i$ is the $i$th column of the weighted matrix square root of $\mathbf P_x$ and $L$ is the dimension of the state space. The total number of sigma points is $2L+1$. The scaling parameter $\lambda$ is defined as 
\begin{equation}\label{eq:sigmapoints3}
	\lambda=\alpha^2( L+\kappa) - L 
\end{equation}
where the constants $\alpha$ determines the spread of the sigma points, $\beta$ incorporates prior knowledge of the distribution of $\mathbf{x}$ and $\kappa$ is a scaling parameter (see ~\cite{Haykin2001} for more details).

The sigma vectors are propagated through the system equations and weighted to form the predicted mean and covariance. The weights are calculated by 
\begin{equation}
	\mathbf W_0^{(m)}=\frac{\lambda}{ L+\lambda} 
\end{equation}
\begin{equation}
	\mathbf W_0^{(c)}=\frac{\lambda}{ L+\lambda}+(1-\alpha^2+\beta) 
\end{equation}
\begin{equation}
	\mathbf W_i^{(m)}=\mathbf W_i^{(c)}=\frac{1}{2( L+\lambda)} \quad i=1, \dots, 2L. 
\end{equation}
Since the observation equation is linear (equation~\ref{ObservationEquation}) the standard Kalman Filter update equations are  used to correct the predicted states. The state estimates from the forward filtering are used to form a new set of sigma points for the smoother, as described above. To compute the smoother gain, the predicted cross-covariance matrix of the states is also calculated. A summary of the URTSS procedure is given in Algorithm~\ref{UKFAlgorithm}. It should be noted that as the disturbance, $e_t\left(\mathbf{r}\right)$ and the measurement noise $ \boldsymbol{\varepsilon}_t$ are additive, therefore the additive form of the URTS is used, otherwise the state vector should be augmented with the noise terms. The additive form reduces the dimension and the number of sigma points used in the algorithm.

\begin{algorithm}
	\begin{small}
	\caption{The Unscented RTS Smoother}\label{UKFAlgorithm} 
	\begin{algorithmic}[1] 
		\State Forward initialisation 
		\begin{equation*}
		 \hat{\mathbf x}_0, \mathbf P_0 
		\end{equation*}
		\State Forward iteration: for $t \in \left\lbrace 0,\cdots, T\right\rbrace $,
		calculate the sigma points $\mathcal X_{i,t}^f$ using \ref{eq:sigmapoints1}-\ref{eq:sigmapoints3} and propogate them through equation \ref{eq:QmatrixForSigmapoints}
% 		\begin{small}
		\begin{equation*}
			\mathcal X_{i,t+1}^{f-}=Q(\mathcal X_{i,t}^f,u_t) 
		\end{equation*}
% 		\end{small}
		calculate the predicted state and the predicted covariance matrix
		\begin{equation*}
			\hat{\mathbf x}_{t+1}^{f-}=\sum_{i=0}^{2L} W_i^{(m)}\mathcal X_{i,t+1}^{f-} 
		\end{equation*}
		\begin{equation*}
			\mathbf P_{t +1}^{f-}=\sum_{i=0}^{2L} W_i^{(c)}(\mathcal X_{i,t+1}^{f-}-\hat{\mathbf x}_{t +1}^{f-})(\mathcal X_{i,t+1}^{f-}-\hat{\mathbf x}_{t +1}^{f-})^\top+\boldsymbol \Sigma_e 
		\end{equation*}
		compute the filter gain, the filtered state and the filtered covariance matrix using the standard Kalman Filter update equations
		\begin{equation*}
			\mathcal K_{t+1}=\mathbf P_{t +1}^{f-}\mathbf C ^\top(\mathbf C \mathbf P_{t +1}^{f-}\mathbf C ^\top+\boldsymbol \Sigma_v)^{-1} 
		\end{equation*}
		\begin{equation*}
			\hat{\mathbf x}_{t+1}^{f}=\hat{\mathbf x}_{t+1}^{f-}+\mathcal K_{t+1}(\mathbf y_{t+1}-\mathbf C\hat{\mathbf x}_{t +1}^{f-}) 
		\end{equation*}
		\begin{equation*}
			\mathbf P_{t+1}^f=(\mathbf I - \mathcal K_{t+1}\mathbf C)\mathbf P_{t +1}^{f-} 
		\end{equation*}
		\State Backward initialisation 
		\begin{equation*}
			\mathbf P_T^b= \mathbf P_T^f, \quad \hat{\mathbf x}^b_T= \hat{\mathbf x}^f_T 
		\end{equation*}
		\State Backward iteration: for $t \in \left\lbrace T-1, \cdots, 0 \right\rbrace $ calculate the sigma points $\mathcal X_{i,t}^b$ and propogate them through equation \ref{eq:QmatrixForSigmapoints}
		\begin{equation*}
			\mathcal X_{i,t+1}^{b-}=Q(\mathcal X_{i,t}^b,u_t) 
		\end{equation*}
		 compute the predicted state, the predicted covariance matrix and the cross-covariance matrix
		\begin{equation*}
			\hat{\mathbf x}_{t+1}^{b-}=\sum_{i=0}^{2L} W_i^{(m)}\mathcal X_{i,t+1}^{b-} 
		\end{equation*}
		\begin{equation*}
			\mathbf P_{t +1}^{b-}=\sum_{i=0}^{2L} W_i^{(c)}(\mathcal X_{i,t+1}^{b-}-\hat{\mathbf x}_{t +1}^{f})(\mathcal X_{i,t+1}^{b-}-\hat{\mathbf x}_{t +1}^{f})^\top+\boldsymbol \Sigma_e 
		\end{equation*}
		\begin{equation*}
			\mathbf M_{t +1}=\sum_{i=0}^{2L} W_i^{(c)}(\mathcal X_{i,t}^{b-}-\hat{\mathbf x}_{t}^{f})(\mathcal X_{i,t+1}^{b-}-\hat{\mathbf x}_{t+1}^{f})^\top 
		\end{equation*}
		 Compute the smoother gain, the smoothed state and the smoothed covariance matrix
		\begin{equation*}
			\mathbf S_t=\mathbf M_{t +1}\left[ \mathbf P_{t +1}^{b-}\right] ^{-1} 
		\end{equation*}
		\begin{equation*}
			\hat{\mathbf x}_t^b=\hat{\mathbf x}_t^f+\mathbf S_t\left[\hat{\mathbf x}_{t+1}^{b}-\hat{\mathbf x}_{t+1}^{b-}\right] 
		\end{equation*}
		\begin{equation*}
			\mathbf P_{t}^{b}=\mathbf P_{t}^{f}+\mathbf S_t\left[\mathbf P_{t+1}^{b}-\mathbf P_{t+1}^{b-} \right]\mathbf S_t^\top 
		\end{equation*}
	\end{algorithmic}
\end{small}
\end{algorithm}

Although the system is nonlinear, the parameters of the system are linear with respect to the state. This is exploited by our procedure where the parameter estimation uses a Least Squares (LS) method that minimises the sum of the squared errors (of a predicted state update) with each new state estimate (see~\ref{LeastSquaresAppendix}).

@@

\section{Results}\label{ResultsSection} The neural field model derived in Section~\ref{NeuralModelSection} was used to generate the data for state and parameter estimation. All parameters for the model are given in Table~\ref{tab:Model Parameters}. An example of the simulated field can be seen in figure~\ref{fig:experimental design}(a). This figure demonstrates the experimental design used: the sensors were placed in the centre of the field (full width at half maximum). This design allowed for uncertainty around the edges of the sensor array, which would be typical from physical data, and alleviated problems associated with boundary conditions. Figure~\ref{fig:experimental design}(b) shows example data from the sensors of the model, and figure~\ref{fig:experimental design}(c) shows the power spectral density (PSD) of the simulated time-series. The PSD shows the typical $1/f$ characteristics of intracranial EEG.

@@

\subsection{Spatial Frequency Analysis} 
Using the model selection techniques described in Section~\ref{SpectralAnalysisSection}, the spatial frequency of the field was used to confirm that the spacing and width of the sensors was adequate to capture the significant dynamics of the field. Figure~\ref{fig:FieldFFT}(a) shows the spatial frequency of the field. The cut-off frequency, $\nu_c$, was taken to be \omg{??} ($-3$~dB point). From equation~\ref{eq:MinimumSensorDistance}, this yielded a minimum sensor separation of \omg{??}. This confirms our separation of \omg{??} was enough to prevent problems associated with aliasing.

The spatial frequency of the observations is shown in Figure~\ref{fig:FreqAnalysis}(c). The peak power of the spatial frequency of the observations was approximately \omg{$??$}~dB. The cut-off frequency (-3~dB point) was taken to correspond to  approximately \omg{$??$}~dB. This provided a conservative estimate of the spatial bandwidth. The oversampling parameter $\rho_b$ was set to $??$ giving a minimum distance between basis functions of $??$ to satisfy Shannon's sampling criterion (from equation~\ref{eq:BasisFunctionSeparation}). \todo{maybe from Parham's derivation... From equation~\ref{} the width was calculated to be... or...}. The width of the basis functions was chosen such that the basis functions had sufficient overlap to represent the dynamic field.

\begin{figure}
	\label{fig:FreqAnalysis} 
	\begin{center}
\subfigure[][]{\includegraphics[scale=0.8]{./Graph/BasisChar.pdf}}
\subfigure[][]{\includegraphics[scale=0.8]{./Graph/SensorChar.pdf}}
\subfigure[][]{\includegraphics[width=1.8in]{./Graph/ObservationFrequencyResponse.pdf}}\\
	\end{center}
	\caption{(a) Field basis function frequency response, (b) Sensor frequency response, (c) The average FFT of the observed field.}

\end{figure}
Following this, the distance between basis function, $\Delta_b$, was set to $2.5$ and the basis function width, $\sigma_{\phi}^2$, was set to 3.6, 3.16 mm full width at half maximum. \wtf{Analysis said 3.9 from Parham's derivation, but 3.6 still overlaps and allows more high frequency details to be represented.}

\subsection{State and Parameter Estimates} 
\label{sec:state_and_param_results}
To demonstrate the performance of the parameter and state estimation, $250$ realisations of the data were generated. Each realisation consisted of 200 ms of data and the estimation was applied to the final 200 ms, allowing the model to stabilise. The initial state and parameters were unknown to the estimator. The distribution of the parameters and the resulting kernel are shown in figure~\ref{fig:Parameters} and figure~\ref{fig:KernelEstimates} respectively. Although a wide distribution can be seen over the parameter estimates the ratio between them remans in a narrow range as shown in figure~\ref{fig:ParametersRatio}. The state estimates for one realization were evaluated by comparing the field reconstruction to the true field using the mean over time of the Root Mean Square Error (RMSE) and standard deviation of the RMSE giving $0.19\pm 0.01$. The RMSE of the field estimate over time is also shown in figure~\ref{fig:RMSE}. One snapshot for the original and the field estimate and their corresponding error are illustrated in figure~\ref{fig:FieldEstimate}.
\begin{figure*}
\subfigure[][]{\includegraphics{./Graph/KernelEstimate_skinny.pdf}}
\subfigure[][]{\includegraphics{./Graph/KernelEstimateFFT_skinny.pdf}}\\
\subfigure[][]{\includegraphics{./Graph/KernelEstimate_wide.pdf}}
\subfigure[][]{\includegraphics{./Graph/KernelEstimateFFT_wide.pdf}}\\
\subfigure[][]{\includegraphics{./Graph/KernelEstimateFFT_skinny_db.pdf}}

@@

\omg{FIRST DESCRIBE AN OVER ALL SUMMARY OF WHAT WE HAVE DONE AND WHAT IS NOVEL}
This paper presents a novel framework for model reduction to facilitate practical state and parameter estimation of neural fields. A basis function decomposition reduces the infinite dimensional neural field (continuous space), to a finite dimensional state-space model. The level of detail of the estimated field is related to the basis functions decomposition, providing a flexible framework where a trade-off can be made between physiological detail and computational complexity. By forming a state-space model, the unscented Rauch Tung Striebel Smoother was implemented for state estimation. Using the sequence of state estimates, a least squares algorithm was implemented for parameter estimation.

\omg{DISCUSS RESULTS: FREQ ANALYSIS. USE THIS TO JUSTIFY KERNEL ERRORS.}
The frequency analysis and model selection tools presented in this paper specify the sensor and field basis function arrangement required to form a state-space model that is capable of capturing the dominant cortical dynamics. These tools can be also used to better design electrophysiological recording techniques to advance model-based neural parameter estimation in order to avoid spatial aliasing. The effect of the basis function decomposition will restrict the frequency content of the estimated field. This is evident in when comparing Figures~\ref{fig:FieldFFT} and~\ref{fig:EstimatedFieldFFT} where the estimated field is spatially band-limited. The spatial frequency content of the field is governed by disturbance covariance and the shape of the connectivity kernel, therefore some information will be lost due to the field decomposition. The implication of this loss of information is the high spatial frequency content of the connectivity kernel. This is demonstrated in Figure~\ref{fig:KernelEstimates} where the frequency content of the estimated connectivity kernel is less than the true kernel.

\omg{DISCUSS THAT EVEN THOUGH THE KERNEL ESTIMATE IS NOT EXACT, IT IS USEFUL TO KNOW THE RATIO OF EXCITATION TO INHIBITION AND THAT RELATIVE MEASURES CAN BE ACCURATELY TRACKED}
Even though the high frequency content of the connectivity kernel is lost through the basis function decomposition, the dominant dynamics are still captured with the estimated reduced model. This is illustrated in~\ref{fig:FieldEstimate}. In addition, the ratio of the kernel parameters follows a tight distribution as seen in Figure~\ref{fig:ParametersRatio}. This indicates that the ratio of excitation to inhibition can be tracked with this approach.

@@

\omg{DISCUSS THE IMPLICATION OF ESTIMATING INTRACORTICAL CONNECTIVITY AND WHERE WE FIT WITHIN OTHER FUNCTIONAL CONNECTIVITY RESEARCH AREAS}
To the authors best knowledge, this is the first paper to propose an estimation framework for intracortical connectivity from electrophysiological data. The importance of this can be illustrated by considering the connectivity structure at this scale can not be estimated using diffusion tensor imaging due to the near isotropic diffusion of water in grey matter~\cite{Assaf2008}. In addition, other non-model based techniques for estimating functional connectivity such as Granger causality \cite{Hesse2003}, the direct transfer function \cite{Kaminski1991}, and partial directed coherence \cite{Sameshima1999} are restricted to the spatial scales of the recording electrodes. The model-based technique proposed in this paper estimates the connectivity structure independently of the recording electrode geometry. An advantage of the auto-regressive (AR) based methods is that they allow for anisotropic connectivity, which is present in cortico-cortical fibres. It should be noted, however, that the AR and the proposed model-based methods are complementary, where connectivity is estimated at different spatial scales. Therefore, we envisage by combining these methods a more realistic model can be acquired. A recent theoretical study~\cite{Jirsa2009} has demonstrated the importance of this multi-scale structure in generating the characteristic rhythms of the brain.

@@

\omg{DISCUSS SHORT-COMINGS AND FUTURE WORK}
In this paper the widths of the connectivity kernel basis functions are assumed to be known. In future work we plan to extend this work to unknown widths. We hope to achieve this by using a sufficiently large set of connectivity basis functions within the physiological plausible limits, and letting the estimation procedure take care of the weights. In addition, we plan to show in future work how the assumption of homogeneity can be relaxed, where heterogeneity can be modelled by careful placement of field basis functions. For example, if cortical regions have a greater density of neural populations, then more basis functions can be used in these areas. In this scenario, model selection and basis function positioning can be achieved using instantaneous frequency analysis of the the neural field or alternative model selection tools based on information criteria. Future work should also be targeted to include a more realistic synaptic response kernel, with finite rise and decay times and finite propagation times into the neural field equations.
